---
title: "HW2"
author: "Jiacheng Wu"
date: "10/6/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

```{r, message=FALSE}
trash_data = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate(sports_balls = round(sports_balls))
```

Read in precipitation from 2018 and 2019

```{r, message=FALSE}
preci18_data = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
             sheet = "2018 Precipitation", 
             range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018)          
```

```{r, message=FALSE}
preci19_data = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
             sheet = "2019 Precipitation", 
             range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2019)          
```
 
Combine rows

```{r, message=FALSE}
preci_data = 
  bind_rows(preci18_data, preci19_data) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month]) 
```

There are `r nrow(trash_data)` valid observations in the Mr. Trash Wheel data frame. There are `r nrow(preci_data)` rows in the 2018&2019 precipitation data frame.
The total precipitation in 2018 is `r preci18_data %>% pull(total) %>% sum`. The total precipitation in 2019 is `r preci19_data %>% pull(total) %>% sum`. The median number of sports balls in a dumpster in 2019 is `r trash_data %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.

## Problem 2

Clean "pols-month.csv"
```{r, message=FALSE}
pols_month_data = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(year = as.numeric(year), month = as.integer(month), month = month.name[month]) %>% 
  mutate(president = ifelse(prez_gop == 0, "dem", "gop")) %>% 
  select(-prez_dem & -prez_gop & -day) %>% 
  relocate(year, month, president)
```

Clean "snp.csv"
```{r, message=FALSE}
 snp_data = 
  read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>%
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day),
         month = month.name[month]) %>% 
  relocate(year, month) %>% 
  arrange(year, month)
```

Tidy "unemployment.csv"
```{r, message=FALSE}
unemployment_data = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
    ) %>% 
  mutate(year = as.numeric(year)) %>% 
  na.omit()
```

Merge the tables
```{r, message=FALSE}
pols_snp_unemployment_data = 
  left_join(pols_month_data, snp_data, by = c("year", "month")) %>% 
  left_join(unemployment_data, by = c("year", "month" ))
```

There are `r nrow(pols_month_data)` observations in `pols-month.csv`. There are `r ncol(pols_month_data)` observations in `pols-month.csv`.
There are `r nrow(snp_data)` observations in `snp.csv`. There are `r ncol(snp_data)` variables in `snp.csv`.
There are `r nrow(unemployment_data)` observations in `unemployment.csv`. There are `r ncol(unemployment_data)` observations in `unemployment.csv`.
After merging the three tables into `pols_snp_unemployment_data`, there are `r nrow(pols_snp_unemployment_data)` observations and and `r ncol(pols_snp_unemployment_data)` variables.



## Problem 3

